### 2023-09-22 article.Rmd

---
title: "Funding programmes for early career researchers influence their independence"
author: "David Janků"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  word_document: default
always_allow_html: yes
bibliography: researcher_independence.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(targets)
library(tidyr)
library(dplyr)
library(cobalt)
library(knitr)
library(psych)
library(kableExtra)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)
library(ez)
library(plm)
library(effectsize)
library(nlme)

```

# Abstract

Independence is important quality constituting one factor of research excellence, often used in assessments for hiring, tenure and promotion in academia [@moherAssessingScientistsHiring2018]. From a bigger picture view, encouraging more independence might also lead to faster exploration of the cognitive map of science and thus lead to faster innovation. Many national research funders have created programmes specifically aimed at increasing independence of early career researchers, yet until recently, there was no quantifiable way to measure changes in independence. We employed recently proposed measure by @vandenbesselaarMeasuringResearcherIndependence2019 to asses whether one such funding programme by Czech Science Foundation is effective at stimulating more independence. We found that funding for early career researchers is effective at stimulating their independence, though the effect is relatively small (2 % of variance) and the source of funding does not matter. The funding also helped selected researchers to catch up with the unfunded group, rather than to outrun them. Additionally, we found differences in levels of independence across disciplines that were stable across time and treatment groups, with Social Sciences and Humanities showing highest independence, while Medical and Heath Sciences and Natural Sciences showing lowest independence.

# Intro

Researchers’ independence from their PhD advisor is one of the criteria often used for assessing researchers for hiring, tenure and promotion (@vanarensbergenDifferentViewsScholarly2014; @moherAssessingScientistsHiring2018), constituting one factor of research excellence.
The promise of independence is that researchers will “challenge the existing common views and open up new lines of research” (@vandenbesselaarMeasuringResearcherIndependence2019, p.2).

An independent investigator is then defined by the US National Academies of Science (NAS) as “one who enjoys independence of thought—the freedom to define the problem of interest and to choose or develop the best strategies and approaches to address that problem–especially during the earlier career phase” (@vandenbesselaarMeasuringResearcherIndependence2019, p.4). Such a definition does not necessarily put independence in relation to other actors (e.g. one’s PhD advisor), but only in relation to one’s own freedom and resources. However other institutions, such as European Research Council via their Starting Grants scheme (@neufeldPeerReviewbasedSelection2013) think of independence in a relational manner, in terms of having one (or a few) publications without the former PhD supervisor being co-author; being PI on a grant; or being lab director or research leader.

Based on some early empirical studies @rojkoScientificPerformanceResearch2022, it seems that researchers in disciplines such as philosophy, psychology; religion, theology; social sciences; the arts, recreation, entertainment, sport; language, linguistics, literature; and geography, biography, history experience significantly higher levels of independence on their PhD advisor than researchers in disciplines such as science and knowledge, organization, computer science, information, documentation, librarianship, institutions, publications; mathematics, natural sciences; and applied sciences, medicine, technology. 
There is also some preliminary evidence that higher independence modestly correlates with number of citations acquired (@rojkoScientificPerformanceResearch2022), but others argue that even the research systems with very low independence can create a lot of scientific impact, and thus the relationships between independence and impact might not be as clear (@glaserIndependenceResearchReview2022)

In recent decades, there has been a worry that researchers’ indepedence is decreasing, illustrated by increase in age at which US (biomedical) researchers received their first grant (NAS, 2005; @danielsGenerationRiskYoung2015), as well as direct measures of new researchers becoming more dependent on their PhD advisors across almost all disciplines in Slovenia (@rojkoScientificPerformanceResearch2022, table 6 on p.9). Perhaps to tackle this trend, many grant agencies decided to start awarding grants for early career researchers, usually up to 8 years after finishing their PhD. Such programmes have often explicitly stated goals of helping researchers become more independent, set up their own research group and open up new lines of research. The examples of such programmes could be Starting Grants by European Research Council (@StartingGrantERC), Junior project and Junior Star by Czech Science Foundation (2020), Walter N. Benjamin Programme by DFG (@WalterBenjaminProgramme), Innovation Research Incentives Scheme by Netherlands Organization of Scientific Research (NWO) (2021)... something from the US would be also good)

Although researcher independence seems to be important feature that is acknowledged by many actors in the research systems and that research funders try to stimulate, there hasn’t been much effort focused on quantifying this phenomenon. Some studies try to measure independence in a simple way, as a ratio of publications the given researcher produced without their supervisor divided by the total number of publications of the given researchers (e.g. @rojkoScientificPerformanceResearch2022). However, such indicator could be misleading in the world of team science, and also only focuses on collaborations, not tracking whether given researcher become independent in their thinking and research topics. [@vandenbesselaarMeasuringResearcherIndependence2019] therefore proposed more complex indicator for measuring researcher independence on the individual level, based on co-authorship network and topical overlaps. Their measure uses only information about publications’ characteristics (authorship and reference list) regardless of it’s impacts, unlike other alternative measures that focus solely on the impact of the independent work in terms of received citations (Dey, et al., 2021, @ALISNovelMetric).

Since the explicit goal of many funding instruments that are being used by research funders all over the world is to help early career researchers become independent, we would like to use the newly created indicators to assess the extent, to which such programmes influence researchers independence.

H1: Researchers, who received funding from programmes aimed at funding early career researchers, will show more independence on their PhD advisor 5 years after receiving the funding than researchers who did not receive this kind of funding.

# Method

## Sample / data

We used the database of all publications produced by researchers employed at Czech research institutions (RIV) combined with a database of all publicly funded research projects and grants in the Czech Republic (CEP), both publicly accessible at https://www.isvavai.cz/, and transformed into a single database with cleaned data, unique person and publication identifiers, and better structure (Hladik, XXX).
This allowed us to identify all researchers who received Junior Grants from the Czech Science Foundation, a funding programme that has an explicit goal to help early career researchers become independent and set up their own research groups. There were 491 projects funded from 2015 to 2020, usually 3 years long (with some being only 2 years long), where the PI must have been up to 8 years after their PhD defence (with potential extension of 2 years for each child the researcher had in that period) and the grant could contribute up to 100 % of their salary (usually, this is limited to 50 % in other funding programmes). From these 491 funded projects, we selected 348 (71 %) grants of the recipients from Czechia, since the remaining recipients were from abroad, and thus would not have their previous publications recorded in the database of researchers employed at Czech research institutions (RIV) that we used for all analyses. There were also 7 cases where one researcher received this grant two times - in these cases, we counted only the first project that was funded, reducing the number of recipients we worked with to 341.

In the second step, we manually searched for PhD advisors of these 341 researchers in the databases of theses and dissertations that cover most universities in Czechia (https://theses.cz and https://dspace.cuni.cz/). We were able to match `r nrow(tar_read(ids_complete))` researchers with a PhD advisor (69,2 % of the Czech subsample). Further, we have not managed to find a suitable matched pair for another `r nrow(tar_read(ids_complete))-nrow(tar_read("matching") %>% filter(treatment == 1) %>% distinct(vedidk))` researchers in the matching procedure, reducing our final sample to `r nrow(tar_read("matching") %>% filter(treatment == 1) %>% distinct(vedidk))`.

For the matched researchers from control groups (see the matching description below in the Analysis section), we didn't search for their supervisors manually, but rather used the "most frequent co-author of the first 5 publications" as a proxy to identify supervisors. Crosschecking this proxy with the actual data about supervisors we have, we found that this algorithm was able to identify supervisors correctly in about 50 % of cases (see the Attachment 1). Using this method, we were able to find `r (nrow(tar_read("final_data") %>% filter(treatment != 1) %>% filter(!is.na(sup_name))) / nrow(tar_read("matching") %>% filter(treatment != 1)))*100` % of the observations in the matched control groups.

These researchers and their PhD advisors with their full publication history (limited to publications created at Czech institutions) created our final sample.

## Measures

*Independence*

We decided to use the recently created independence indicator by @vandenbesselaarMeasuringResearcherIndependence2019, since it is based on scientific production (rather than impact) and captures the phenomena in the most complete manner. The indicator has 4 parts: I1: The eigenvector centrality of the supervisor in the researcher’s co-author network; I2: The clustering coefficient of the supervisor in the researcher’s co-author network; I3: The share of own papers of the researcher Pns/Pall where Pns is the number of publications of the researcher, not coauthored with the former supervisor(s) and Pall is the total number of publications of the researcher; and I4: The share of own research topics of the researcher Tns/Tall where Tns is the number of research topics of the researcher in which the former supervisor(s) is not active, and Tall is the total number of topics of the researcher.

However, we have made some changes to the indicator. After consulting with its authors, we have changed the aggregation formula to 
$$RII = ((1- I1) + I2 + I3 + I4) / 4$$ 
since the previous formula had a small error (it weighted I4 double,, i.e. 2*I4). In determining the proportion of a researcher's own research topics (I4), we opted for text-based topic modeling utilizing publication titles and abstracts, diverging from the originally suggested bibliographic coupling method. Initially, we conducted a random sampling of 1200 publications from a set of all publications produced by researchers affiliated with Czech institutions as cataloged in the RIV database, estimated number of topics using Griffiths2004 indicator and calculated LDA topic model on this set of publications. Then we  compiled all publications generated by the researchers within our sample and their respective supervisors. We then calculated the posterior distribution of topics for this compilation, employing a pre-trained topic model. This way we wanted to mitigate the endogeneity of the topic modelling which would be stronger if we calculated topic model directly on the publications of our sample researchers.

<!-- Finally, the original indicator, while tracking not only collaboration independence but also topical independence, is still skewed significantly towards collaboration independence (three-quarters of the indicator measure collaboration, while only one-quarter measures topics). Since we consider cognitive/topical independence an important asset of this indicator, we have decided to adjust the formula a bit further and drop the I3 part, to give more weight to the cognitive/topical independence in the whole indicator. Throughout this paper, we will thus refer to the original indicator as RIIo, our adjusted version as RIIa and the part of the indicator measuring only the cognitive independence RIIc. -->

*Career age*

We measured researchers’ career age as the number of years since their first publication.

*Discipline*

We used the highest level of the OECD classification of disciplines, containing categories Natural Sciences; Engineering and Technology; Medical and Health Sciences; Agricultural Sciences; Social Sciences; Humanities. We chose this classification because it is widely used and also used in the dataset we worked with. For the matching we used the 42 sub-categories of this classification to achieve more granularity.

We attributed a discipline to a given researcher based on disciplines attributed to their each publication, choosing the most frequent discipline for each author.

*Interdisciplinary proportion*

This measure was calculated as the ratio of publications authored by the given researcher which were not assigned to the author’s main discipline.

*Number of publications*

We counted all publications the given researcher produced before the intervention year (i.e. receiving a Junior Grant) while affiliated with the Czech institution (our database only contained publications affiliated with Czech institutions). We only counted publications of types: journal articles, conference proceedings, book chapters, and full books.

*Number of publications in Web of Science and Scopus*

We counted all publications the given researcher produced before the intervention year (i.e. receiving a Junior Grant) while affiliated with the Czech institution (our database only contained publications affiliated with Czech institutions), that were also linked in Web of Science or Scopus (which means they were likely published in more recognised journals). We only counted publications of types: journal articles, conference proceedings, book chapters, and full books.

*Number of grants received prior to Junior Grant*

We calculated how many (Czech) grants each researcher has received prior to the intervention year (i.e. receiving a Junior Grant).

*Career age when receiving the first grant*

We calculated the career age of each researcher when they received their first grant.

*Gender*

We used Genderize.io API to estimate the probability of perceived gender based on the first names of researchers. Where available, information about the nationality of researchers was included in the API call to improve accuracy. 
We have also manually specified gender for 2 obrsevtation from the treatment group for which the gender was not calculated automatically.

## Analysis

We analysed the data in the R version 4.2.3. using R studio and [this code](https://github.com/david-janku/juniors). For the construction of the independence indicator, we used network modelling (maybe more specific?) and text-based topic modelling (maybe more specific??).

For the construction of the control group, we implemented propensity matching using the package MatchIt (citation?) and matching exactly by the discipline (42 sub-categories of OECD classification) and treatment year, and approximately by number of publications in total, number of publications in WoS/Scopus, career age, interdisciplinary proportion, number of grants received prior to Junior Grant, career age when receiving the first grant and gender. All of these variables were calculated at the intervention year (i.e. year when the given researcher from experimental group received the Junior Grant). The distance was set to "mahalanobis". The matching was done in 1:3 ratio and with replacement. Manual search for supervisors of the matched control group observations was done by ranking the control group observation by distance from the treatment group observation and first searching for the supervisor of the closest control group observation, and if it was not possible to find it, then repeat the same for the second closest, etc... In the end each treatment observation was matched with exactly one control group observation which was closest but at the same time for which it was possible to identify their supervisor.

We have created 2 control groups: one using the matching above and one similar but specifically containing only researchers who received their first grant (other than Junior Grant) around the same career age as researchers in our intervention group. That allowed us to better track the impact of any grant funding vs specifically Junior Grants funding on researchers’ independence.

The difference between groups was tested using paired t-test and subsequent differences between discipline groups was tested using ANOVA post-hoc Tukey test.

# Results

## Descriptives

### Missing data
```{r echo=FALSE}
d <- tar_read(full_indicator)
d$disc_ford <- as.factor(d$disc_ford)

na_counts <- sum(is.na(d$RII))
na_percentage <- round((na_counts / nrow(d))*100, 1)

na_sup <- round((sum(is.na(d$sup_name))/na_counts)*100, 1)
na_pubs <- round((nrow(d %>% filter(!is.na(sup_name)) %>% filter(is.na(pubs_number)))/na_counts)*100, 1)

na_pubs_after <- round((nrow(d %>% filter(!is.na(sup_name)) %>% filter(is.na(pubs_number)) %>% filter(independence_timing == "after_intervention"))/nrow(d %>% filter(!is.na(sup_name)) %>% filter(is.na(pubs_number))))*100, 1)

na_treatment <- nrow(d %>% filter(is.na(RII)) %>% filter(treatment == 1))
na_control1 <- nrow(d %>% filter(is.na(RII)) %>% filter(treatment == 0))
na_control2 <- nrow(d %>% filter(is.na(RII)) %>% filter(treatment == 2))

# s <- d %>% filter(is.na(RII))
#                   
# f_after <- d %>% filter(!is.na(RII)) %>% filter(independence_timing == "after_intervention") 
# f_before <- d %>% filter(!is.na(RII)) %>% filter(independence_timing == "before_intervention") 
# 
# 
# v_after <- c(f_after$vedidk[f_after$treatment == 1 ], NA)
# v_before <- c(f_before$vedidk[f_before$treatment == 1 ], NA)
# 
# r_after <- f_after %>% filter(vedidk_treatment %in% v_after)
# r_before <- f_before %>% filter(vedidk_treatment %in% v_before)
# 
# 
# 
# v_after_control <- c(r_after$vedidk[r_after$treatment != 1 ], r_after$vedidk_treatment)
# v_before_control <- c(r_before$vedidk[r_before$treatment != 1 ], r_before$vedidk_treatment)
# 
# r_after_final <- r_after %>% filter(vedidk %in% v_after_control)
# r_before_final <- r_before %>% filter(vedidk %in% v_before_control)

# r_test <- r_after %>% filter(treatment != 2)
# v_after_test <- c(r_test$vedidk[r_test$treatment != 1 ], r_test$vedidk_treatment)
# r_test_final <- r_test %>% filter(vedidk %in% v_after_test)

# final_before_control1 <- nrow(r_before_final %>% filter(treatment == 0)) 
# final_before_control2 <- nrow(r_before_final %>% filter(treatment == 2))
# 
# 
# final_after_control1 <- nrow(r_after_final %>% filter(treatment == 0)) 
# final_after_control2 <- nrow(r_after_final %>% filter(treatment == 2))

# a <- intersect(
#     r_after_final$vedidk,
#     r_before_final$vedidk
# )
# 
# 
# all <- rbind(r_before_final, r_after_final) %>% filter(vedidk %in% a)

filtered_by_subclass <- d %>%
     group_by(subclass, independence_timing) %>%
     filter(!any(is.na(RII))) %>%
     ungroup()

filtered_subclass <- nrow(filtered_by_subclass %>% distinct(subclass))
subclass_all <- nrow(d %>% distinct(subclass))
filtered_subclass_perc <- round((filtered_subclass/subclass_all)*100, 1)

final_before_control1 <- nrow(filtered_by_subclass %>% filter(independence_timing == "before_intervention") %>% filter(treatment == 0))
final_before_control2 <- nrow(filtered_by_subclass %>% filter(independence_timing == "before_intervention") %>% filter(treatment == 2))

final_after_control1 <- nrow(filtered_by_subclass %>% filter(independence_timing == "after_intervention") %>% filter(treatment == 0))
final_after_control2 <- nrow(filtered_by_subclass %>% filter(independence_timing == "after_intervention") %>% filter(treatment == 2))

```

The independence indicator was not possible to calculate for `r na_counts` (`r na_percentage` % of observations). This is in `r na_sup` % caused by missing data about supervisors, and in `r na_pubs` % due to given researcher not having any publications in the given period (`r na_pubs_after` % of these cases are in the period after the intervention, which sometimes mean in the past few years).

There are `r na_treatment` missing observations in the treatment group, `r na_control1` in the unfunded control group and  `r na_control2` in the funded control group.
If we only keep observations that have a match with non-missing values, we are left with a sample of `r final_before_control1` matched observations in unfunded control group and `r final_before_control2` matched observations in funded control group before intervention and a sample of `r final_after_control1` matched observations in unfunded control group and `r final_after_control2` matched observations in funded control group after intervention.

When Looking at disciplines, we see that there are slightly more missing data in Agricultural and veterinary sciences and the Humanities and the Arts.

```{r echo=FALSE}

dt <- d
dt$missing <- ifelse(is.na(dt$RII), "Missing", "Complete")

# Calculate counts and proportions
df_summary <- dt %>%
  group_by(field, missing) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(Proportion = count / sum(count)) %>%
  pivot_wider(names_from = missing, values_from = c(count, Proportion))

# If there are any NA values in the 'count_Missing' column, replace them with 0
df_summary$count_Missing[is.na(df_summary$count_Missing)] <- 0

# Calculate Missing / (Missing + Complete) in percentages for each disc_ford
df_summary <- df_summary %>%
  mutate(Proportion = ifelse(!is.na(`count_Missing`) & !is.na(`count_Complete`), `count_Missing` / (`count_Missing` + `count_Complete`) * 100, NA))

# If there are any NA values in the 'Proportion' column, replace them with 0
df_summary$Proportion[is.na(df_summary$Proportion)] <- 0

# Remove unnecessary columns and rename the others
df_summary <- df_summary %>%
  select(field, count_Complete, count_Missing, Proportion) %>%
  rename(
    Complete = count_Complete,
    Missing = count_Missing,
    Proportion_Missing = Proportion
  )

# Print the updated table
kable(df_summary, digits = 3, caption = "Counts and Proportions of Missing data by disc_ford") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

### Before intervention

POZN: here we show data

```{r echo=FALSE}

df <- filtered_by_subclass %>% filter(independence_timing == "before_intervention")
numeric_vars <- df %>% 
  select(where(is.numeric)) %>% 
  select(-weights, -funded_control, -id, -treatment_year, -career_start_year) %>% 
  names()

# descriptive stats for each treatment level
table_before_1 <- df %>% filter(treatment == 1) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)
table_before_0 <- df %>% filter(treatment == 0) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)
table_before_2 <- df %>% filter(treatment == 2) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)

colnames(table_before_0) <- paste0(colnames(table_before_0), "_0")
colnames(table_before_2) <- paste0(colnames(table_before_2), "_2")

combined_table_before <- cbind(table_before_1, table_before_0[, -1], table_before_2[, -1])

kable(combined_table_before, 
      digits = 3, align = "c", booktabs = TRUE) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  add_header_above(c(" " = 1, "Treatment Group" = 5, "Unfunded Control Group" = 4, "Funded Control Group" = 4))

```

### After intervention

```{r echo=FALSE}

df <- filtered_by_subclass %>% filter(independence_timing == "after_intervention")
numeric_vars <- df %>% 
  select(where(is.numeric)) %>% 
  select(-weights, -funded_control, -id, -treatment_year, -career_start_year) %>% 
  names()

# descriptive stats for each treatment level
table_after_1 <- df %>% filter(treatment == 1) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)
table_after_0 <- df %>% filter(treatment == 0) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)
table_after_2 <- df %>% filter(treatment == 2) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)

colnames(table_after_0) <- paste0(colnames(table_after_0), "_0")
colnames(table_after_2) <- paste0(colnames(table_after_2), "_2")

combined_table_after <- cbind(table_after_1, table_after_0[, -1], table_after_2[, -1])

kable(combined_table_after, 
      digits = 3, align = "c", booktabs = TRUE) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  add_header_above(c(" " = 1, "Treatment Group" = 5, "Unfunded Control Group" = 4, "Funded Control Group" = 4))

```

## Matching

```{r echo=FALSE}

removed_treatment_unfunded <- (tar_read(ids_complete) %>% distinct(vedidk) %>% nrow())-(tar_read(matching) %>% group_by(subclass) %>% filter(!any(treatment == 2)) %>%
  ungroup() %>% filter(treatment == "1") %>% distinct(vedidk) %>% nrow())

removed_treatment_funded <- (tar_read(ids_complete) %>% distinct(vedidk) %>% nrow())-(tar_read(matching) %>% group_by(subclass) %>% filter(!any(treatment == 0)) %>%
  ungroup() %>% filter(treatment == "1") %>% distinct(vedidk) %>% nrow())

unique_treatment_unfunded <- d %>% group_by(subclass) %>% filter(!any(treatment == 2)) %>% ungroup() %>% filter(treatment == "1") %>% distinct(vedidk) %>% nrow()

unique_treatment_funded <- d %>% group_by(subclass) %>% filter(!any(treatment == 0)) %>%
  ungroup() %>% filter(treatment == "1") %>% distinct(vedidk) %>% nrow()

unique_control_unfunded <- d %>% filter(treatment == 0) %>% distinct(vedidk) %>% nrow()

unique_control_funded <- d %>% filter(treatment == 2) %>% distinct(vedidk) %>% nrow()

replaced_control_unfunded <- unique_treatment_unfunded-unique_control_unfunded

replaced_control_funded <- unique_treatment_funded-unique_control_funded

overlap_control <- unique_control_unfunded+unique_control_funded-(d %>% filter(treatment != 1) %>% distinct(vedidk) %>% nrow())

```

After matching, all standardized mean differences for the covariates were below 0.1 and all standardized mean differences for squares and two-way interactions between covariates were below .15, indicating adequate balance.

Matching process removed `r removed_treatment_unfunded` treatment observations from matching with unfunded control group and `r removed_treatment_funded` treatment observations from matching with funded control group. This observation was removed because the researchers had no publications prior to the year of application for the grant, which would make it not possible to calculate the independence score for them anyway.

The resulting sample had `r unique_treatment_unfunded` unique treatment group observations paired with `r unique_control_unfunded` unique unfunded control group observations suggesting that `r replaced_control_unfunded` observations from unfunded control group were matched to multiple treatment group observations. Similarly, there were `r unique_treatment_funded` unique treatment group observations paired with `r unique_control_funded` unique funded control group observations, suggesting that `r replaced_control_funded` funded control group observations were matched to multiple treatment group observations.

BTW: there is an overlap between control group observations: `r overlap_control` observations from the unfunded control group also appears in the funded control group

```{r echo=FALSE}
# # Create a subset for treatment group and control group 1
# subset_1 <- (tar_read(matching)) %>% filter(treatment != 2)
# subset_2 <- (tar_read(matching)) %>% filter(treatment != 0)
# 
# # Generate balance table for treatment vs control group 1
# balance_results_1 <- bal.tab(treatment ~ career_lenght + pubs_total + ws_pubs + interdisc_proportion + grants + gender, data = subset_1, disp = c("mean", "std"))
# balance_results_2 <- bal.tab(treatment ~ career_lenght + pubs_total + ws_pubs + interdisc_proportion + grants + gender, data = subset_2, disp = c("mean", "std"))
# 
# # Define a function to manually extract balance statistics from bal.tab object
# extract_balance_stats <- function(bal_results) {
#   covariates <- rownames(bal_results$Balance)
#   means_treat <- bal_results$Balance$`M.1.Un`
#   means_ctrl <- bal_results$Balance$`M.0.Un`
#   diff <- bal_results$Balance$`Diff.Un`
#   
#   data.frame(Covariate = covariates, 
#              Mean_Treatment = means_treat, 
#              Mean_Control = means_ctrl, 
#              Diff = diff)
# }
# 
# # Extract balance stats for subset_1 and subset_2
# balance_stats_1 <- extract_balance_stats(balance_results_1)
# balance_stats_2 <- extract_balance_stats(balance_results_2)
# 
# # Print tables in R Markdown
# knitr::kable(balance_stats_1, caption = "Balance Statistics: Treatment vs Control Group 1") %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
# knitr::kable(balance_stats_2, caption = "Balance Statistics: Treatment vs Control Group 2") %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

##Unfunded group

unf_sum <- summary(tar_read(matched_obj_unfunded), un = TRUE)

# summary(tar_read(matched_obj_unfunded))
#     
# plot(unf_sum, var.order = "unmatched")

aa <- cobalt::bal.tab(tar_read(matched_obj_unfunded), un = TRUE, stats = c("m", "v", "ks"), binary = "std")

# aa <- aa[!grepl("^disc_ford", names(aa))]

# aa <- aa[!grepl("^disc_ford", names(aa$Balance))]

bal_df <- aa$Balance

bal_df <- bal_df[!grepl("^disc_ford", rownames(bal_df)), ]

# Display the cleaned-up table with kable
kable(bal_df, "pipe", caption = "Balance Statistics For Unfunded Control Group") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

aa$Balance <- bal_df

cobalt::love.plot(aa, binary = "std", var.order = "adjusted")

##Funded group

fun_sum <- summary(tar_read(matched_obj_funded), un = TRUE)

# summary(tar_read(matched_obj_funded))
#     
# plot(fun_sum, var.order = "unmatched")

bb <- cobalt::bal.tab(tar_read(matched_obj_funded), un = TRUE, stats = c("m", "v", "ks"), binary = "std")

# aa <- aa[!grepl("^disc_ford", names(aa))]

# aa <- aa[!grepl("^disc_ford", names(aa$Balance))]

bal_df_funded <- bb$Balance

bal_df_funded <- bal_df_funded[!grepl("^disc_ford", rownames(bal_df_funded)), ]

# Display the cleaned-up table with kable
kable(bal_df_funded, "pipe", caption = "Balance Statistics For Funded Control Group") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

bb$Balance <- bal_df_funded

cobalt::love.plot(bb, binary = "std", var.order = "adjusted")

```

## Main results

To ascertain the differences in independence scores before and after intervention, while accounting for the treatment type, academic disciplines, and paired nature of the data, we used repeated measures ANOVA. This allowed us to measure the main effect of change in independence across time, across disciplines, and the main treatment effect was the interaction between treatment type and time.

<!-- POZN: only for comparison: -->
<!-- To estimate the treatment effect and its standard error, we fit a linear regression model with 1978 earnings as the outcome and the treatment, covariates, and their interaction as predictors and included the full matching weights in the estimation. The lm() function was used to fit the outcome, and the comparisons() function in the marginaleffects package was used to perform g-computation in the matched sample to estimate the ATT. A cluster-robust variance was used to estimate its standard error with matching stratum membership as the clustering variable. -->

<!-- The estimated effect was XXXX (SE = XXXX, p = XXXX), indicating that the average effect of the treatment for those who received it is to increase earnings. -->

```{r echo=FALSE}

pp <- d %>%
     group_by(subclass) %>%
     filter(!any(is.na(RII))) %>%
     ungroup()

pp <- pp %>%
     group_by(vedidk, subclass) %>%  # Assuming 'id' is your current unique identifier
     mutate(subject_id = cur_group_id()) %>%  # This creates a unique ID for each group
     ungroup()

diff_data <- pp %>%
  select(subject_id, independence_timing, RII) %>%
  spread(independence_timing, RII) %>%
  mutate(RII_diff = after_intervention - before_intervention) %>%
  select(subject_id, RII_diff)

pp <- left_join(pp, diff_data, by = "subject_id")

filtered_by_subclass_1 <- pp %>% 
    group_by(subclass) %>% 
    filter(!any(treatment == 2)) %>% 
    ungroup() %>%
    mutate(treatment = factor(treatment, levels = c(0, 1)))

filtered_by_subclass_1$subclass <- as.factor(filtered_by_subclass_1$subclass)
filtered_by_subclass_1$treatment <- as.factor(filtered_by_subclass_1$treatment)
filtered_by_subclass_1$independence_timing <- as.factor(filtered_by_subclass_1$independence_timing)
filtered_by_subclass_1$field <- as.factor(filtered_by_subclass_1$field)

# this DiD model unfortunately doesnt work for some reason that I cannot resolve, but it is possible to calculate the same using repeated measures anova or mixes effects model, so it should be fine

# # Run regression model with interaction term
# did_model <- plm::plm(RII ~ independence_timing * treatment + factor(subclass),
#                  data = filtered_by_subclass_1,
#                  index = c("subclass"),
#                  model = "within")
# 
# # Calculate clustered standard errors
# clustered_se <- coeftest(did_model, vcov = function(x) vcovHC(x, type = "HC1", cluster = "subclass"))

## here should be some assumptions tests that I will run to see whtehr it is correct, but later I will comment it so that it doesnt show in the final paper

# model <- lmer(RII ~ treatment * independence_timing + (1 | subclass), data = filtered_by_subclass_1)
# # summary(model)
# # r.squaredGLMM(model)
# 
# model_no_interaction <- lmer(RII ~ treatment + independence_timing + (1 | subclass), data = filtered_by_subclass_1)
# 
# RSS_no_interaction <- sum(residuals(model_no_interaction)^2)
# RSS_full <- sum(residuals(model)^2)
# delta_RSS <- RSS_no_interaction - RSS_full
# R2_interaction <- delta_RSS / RSS_no_interaction
# # print(R2_interaction)
# 
# 
# model_disc <- lmer(RII ~ treatment * independence_timing * field + 
#               (1 + independence_timing|subclass), 
#               data = filtered_by_subclass_1)

# summary(model_disc)

##checking assumptions fro repeated measures ANOVA

result <- aov(RII ~ treatment * independence_timing + Error(subclass/independence_timing), data=filtered_by_subclass_1)
# summary(result)
eta_sq_result <- eta_squared(result)

# Normality: You can check the normality of the residuals using a Q-Q plot and the Shapiro-Wilk test.

# Check residuals from lm
# plot(resid(result))

#Linearity and Homoscedasticity (Equal Variances) of Residuals: If the assumptions hold, you should not see any obvious patterns or funnel shapes in the plot.

# plot(result$fitted.values, resid(result), 
#      xlab = "Fitted values", ylab = "Residuals", 
#      main = "Residuals vs. Fitted Values")
# abline(h = 0, col = "red")

# Normality of Residuals: 
# qqnorm(resid(result))
# qqline(resid(result))
# 
# shapiro.test(resid(result))

# Independence of Residuals: If your data has a time sequence (e.g., repeated measurements over time), then you would be concerned about the independence of residuals. For this, you can use the Durbin-Watson test from the car package. The Durbin-Watson test statistic ranges from 0 to 4. A value around 2 suggests no autocorrelation, while values below 1 and above 3 are cause for concern.

# library(car)
# durbinWatsonTest(result)

# Multicollinearity: Check the variance inflation factor (VIF) for each predictor. Typically, a VIF above 5-10 indicates multicollinearity.

# vif(result)

# Outliers and Influence Points: Leverage vs. studentized residuals plot can be used to identify influential points.

# plot(hatvalues(result), rstudent(result),
#      xlab = "Leverage", ylab = "Studentized Residuals",
#      main = "Residuals vs. Leverage")
# abline(h = c(-2,2), col = "red", lty = 2)

# Sphericity: Use the mauchly.test function on your model. If the test is significant, then the assumption of sphericity has been violated.

# mauchly.test(model_aov)

# Homogeneity of Variance: Use Levene's Test. The car package offers leveneTest.

# library(car)
# leveneTest(RII ~ treatment, data = filtered_by_subclass_1)

# Outliers:You can inspect boxplots or use Mahalanobis distance.

# boxplot(RII ~ treatment, data = filtered_by_subclass_1)

##conducting repeated measures ANOVA

# filtered_by_subclass_1 <- filtered_by_subclass_1 %>%
#      group_by(vedidk, subclass) %>%  # Assuming 'id' is your current unique identifier
#      mutate(subject_id = cur_group_id()) %>%  # This creates a unique ID for each group
#      ungroup()

# #this is preferable because it takes into account the matched nature of the data, as well as repeated measures nature of the data -> see chatGPT conversation "ANOVA `wid` parameter clarification"
# result <- aov(RII ~ treatment * independence_timing + Error(subclass/independence_timing), data=filtered_by_subclass_1)
# # summary(result)
# eta_sq_result <- eta_squared(result)

#this is less preferable because it takes into account only the repeated measure nature of the data, not the matched nature -> you can see that when you run this tests, there is warning that says "there were non-unique subclass values across the between-subjects variable (treatment). Essentially, ezANOVA detected the problem and internally "fixed" it by making subclass unique across the treatment groups, so it behaves similarly to subject_id you later created." -> this means that it effectively converts the subclass to subject_id anyway
# result <- ezANOVA(data = filtered_by_subclass_1,
#                  dv = RII,
#                  wid = subclass,
#                  within = .(independence_timing),
#                  between = treatment)
# # print(result)

## adding disciplines:

filtered_data_disc_1 <- filtered_by_subclass_1 %>%
  group_by(field) %>%
  filter(n() >= 10)

filtered_data_disc_1$field <- droplevels(filtered_data_disc_1$field)

# Calculate weights based on group size (you can adjust this based on other criteria if needed)
group_sizes <- table(filtered_data_disc_1$field)
max_size <- max(group_sizes)
filtered_data_disc_1$weights <- max_size / group_sizes[filtered_data_disc_1$field]

#this model addresses both unequal group sizes (via weights) and heteroscedasticity (via varIdent). Additionally, I've added a first-order autoregressive structure (corAR1) assuming you have repeated measures on the same subjects and to account for their paired nature (indicated by subclass/independence_timing)
#this model directly tests whether there is a difference in disciplines - anyway it tests each discipline agains each other, so it doesnt produce the neat result of "yes, there are some significant differences between disciplines", but more nuanced version of "there is specifically difference between this and this discipline"

model_gls_weighted <- gls(RII ~ field, data = filtered_data_disc_1, 
                          weights = varIdent(form = ~ 1 | field),
                          method = "ML",
                          correlation = corAR1(form = ~ 1 | subject_id))

# summary(model_gls_weighted)

#this model below is the same as the above but with added predictors

model_gls_interactions_subclass <- gls(RII ~ field * treatment * independence_timing, 
                              data = filtered_data_disc_1, 
                              weights = varIdent(form = ~ 1 | field),
                              method = "ML",
                              correlation = corAR1(form = ~ 1 | subclass/independence_timing))

# aa <- summary(model_gls_interactions_subclass)

#one potential strategy to get the neat result of "yes, there are some significant differences between disciplines" is following: create two similar models

model_without_field <- gls(RII ~ treatment + independence_timing + treatment:independence_timing, 
                           data = filtered_data_disc_1, 
                           correlation = corCompSymm(form = ~ 1 | subclass/independence_timing), 
                           method = "ML")

model_with_field_main_effect <- gls(RII ~ treatment + field + independence_timing + treatment:independence_timing, 
                                    data = filtered_data_disc_1, 
                                    correlation = corCompSymm(form = ~ 1 | subclass/independence_timing), 
                                    method = "ML")

anova_result <- anova(model_without_field, model_with_field_main_effect)

model_summary <- summary(model_with_field_main_effect)

# Your result specifically shows that the model_with_field_main_effect is significantly better at explaining the variance in your response variable compared to the model_without_field. The p-value is extremely small (1e-04), indicating a significant main effect of "field".
# 
# In other words, the main effect of "field" significantly improves the model fit, suggesting that the variable "field" has a significant main effect on the response variable RII after accounting for other predictors in the model.

#this model does NOT address unequal group sizes (via weights) and heteroscedasticity

# result_disc <- ezANOVA(data = filtered_data_disc_1,
#                  dv = RII,
#                  wid = subclass,
#                  within = .(independence_timing),
#                  between = .(treatment, field))
# # print(result_disc)

##here I calculate the interaction of filed * treatment not from the full RII score, but from RR delta, ie change in RII from before to after intervetion. below the results are plotted, showing that the effect of treatment on disciplines is universal, expect for social sciences where there is almost no effect. it also shows that almost in all cases the effect of time is positive - i.e people are getting more independent over time, in some case by quite a lot (e.g. 0.2 in RII, which is 20 % oif the RII score)

# filtered_data_disc_1 <- filtered_data_disc_1 %>% 
#     filter(independence_timing != "after_intervention")
# 
# means_table_disc_1 <- filtered_data_disc_1 %>%
#      group_by(field, treatment) %>%
#      summarise(mean_RII = mean(RII_diff, na.rm = TRUE))
# 
# # Set up sum coding for "field"
# contrasts(filtered_data_disc_1$field) <- contr.sum(levels(filtered_data_disc_1$field))
# 
# 
# model_wls <- lm(RII_diff ~ field * treatment, data = filtered_data_disc_1, weights = weights)
# summary(model_wls)
# 
# 
# contrasts(filtered_data_disc_1$field)
# 
# # Extract the variance-covariance matrix
# vcov_mat <- vcov(model_wls_sum_coded)
# 
# # Coefficient for Social Sciences
# coef_SS <- -sum(coef(model_wls_sum_coded)[c("field1", "field2", "field3")])
# 
# # Variance for Social Sciences
# var_SS <- sum(diag(vcov_mat[2:4, 2:4])) + 2*sum(vcov_mat[2,3:4]) + 2*sum(vcov_mat[3,4])
# 
# # Standard Error for Social Sciences
# se_SS <- sqrt(var_SS)
# 
# # t-value for Social Sciences
# t_value_SS <- coef_SS / se_SS
# 
# # Degrees of Freedom
# df <- df.residual(model_wls_sum_coded)
# 
# # p-value for Social Sciences
# p_value_SS <- 2 * pt(-abs(t_value_SS), df)
# 
# coef_SS
# se_SS
# t_value_SS
# p_value_SS
# 
# vcov_matrix <- vcov(model_wls_sum_coded)
# 
# 
# var_interaction <- vcov_matrix["treatment1", "treatment1"] + 
#                    vcov_matrix["field1:treatment1", "field1:treatment1"] + 
#                    vcov_matrix["field2:treatment1", "field2:treatment1"] + 
#                    vcov_matrix["field3:treatment1", "field3:treatment1"] + 
#                    2*(vcov_matrix["treatment1", "field1:treatment1"] + 
#                       vcov_matrix["treatment1", "field2:treatment1"] + 
#                       vcov_matrix["treatment1", "field3:treatment1"])
# 
# se_interaction <- sqrt(var_interaction)
# 
# t_value <- -0.01905 / se_interaction
# 
# df_residual <- df.residual(model_wls_sum_coded)
# p_value <- 2 * (1 - pt(abs(t_value), df_residual))

## visualise field*treatment

# means_table <- filtered_by_subclass_1 %>%
#   filter(independence_timing != "before_intervention") %>% 
#   group_by(field, treatment) %>%
#   summarise(mean_RII = mean(RII_diff, na.rm = TRUE))
# 
# sd_counts <- filtered_by_subclass_1 %>%
#   filter(independence_timing != "before_intervention") %>% 
#   group_by(field, treatment) %>%
#   summarise(sd_RII = sd(RII_diff, na.rm = TRUE),
#             n = n())
# 
# means_table <- merge(means_table, sd_counts, by = c("field", "treatment")) %>%
#   mutate(se_RII = sd_RII / sqrt(n))
# 
# 
# 
# 
# 
# ggplot(means_table, aes(x=field, y=mean_RII, fill=treatment)) +
#     geom_bar(stat="identity", position=position_dodge(width=0.8), width=0.7) +
#     geom_errorbar(aes(ymin=mean_RII-se_RII, ymax=mean_RII+se_RII),
#                   position=position_dodge(width=0.8), width=0.3) +
#     geom_text(aes(label=n, y= 0.5 * mean_RII), # position in the middle
#               position=position_dodge(width=0.8), vjust=0.5) + # centered vertically
#     theme_minimal() +
#     labs(y="Mean RII Difference", x="Field", fill="Treatment", title="Comparison of differences in RII across fields and treatment group") +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

In the first analysis, we examined the effects of treatment, independence_timing, and their interaction on the dependent variable, researcher independence (RII) for the treatment group and unfunded control group.

There was a significant main effect of independence_timing on RII, F(`r summary(result)$"Error: subclass:independence_timing"[[1]]$"Df"[1]`, `r format(summary(result)$"Error: subclass:independence_timing"[[1]]$"Df"[2])`) = `r format(summary(result)$"Error: subclass:independence_timing"[[1]]$"F value"[1], digits=2)`, p = `r format(summary(result)$"Error: subclass:independence_timing"[[1]]$"Pr(>F)"[1], digits=4)`). This effect accounted for approximately `r format(eta_sq_result[eta_sq_result$Parameter == "independence_timing", "Eta2_partial"] * 100, digits=2)`% of the observed variance (partial \(\eta^2 = `r eta_sq_result[eta_sq_result$Parameter == "independence_timing", "Eta2_partial"]`\), 95% CI [`r eta_sq_result[eta_sq_result$Parameter == "independence_timing", "CI_low"]`, `r eta_sq_result[eta_sq_result$Parameter == "independence_timing", "CI_high"]`]).

Importantly, there was a significant interaction effect between treatment and independence_timing on independence, F(`r format(summary(result)$"Error: Within"[[1]]$"Df"[2])`, `r format(summary(result)$"Error: Within"[[1]]$"Df"[3])` = `r format(summary(result)$"Error: Within"[[1]]$"F value"[2], digits=2)`, p = `r format(summary(result)$"Error: Within"[[1]]$"Pr(>F)"[2], digits=2)`. This interaction effect explained about `r format(eta_sq_result[eta_sq_result$Parameter == "treatment:independence_timing", "Eta2_partial"] * 100, digits=2)`% of the variability (partial \(\eta^2 = `r eta_sq_result[eta_sq_result$Parameter == "treatment:independence_timing", "Eta2_partial"]`\), 95% CI [`r eta_sq_result[eta_sq_result$Parameter == "treatment:independence_timing", "CI_low"]`, `r eta_sq_result[eta_sq_result$Parameter == "treatment:independence_timing", "CI_high"]`]).

To assess effects of study field on researcher independence, we chose GLS weighted model to account for unequal group sizes and heteroscedastic variances, while keeping the repeated measures design. First, we compared models with and without the field main effect and observed a notable improvement in model fit for the model where main effect of field was included. This suggests that there are statistically significant differences in the RII based on the field of study. Specifically, adding the field of study to the model resulted in a significant increase in the log-likelihood by `r anova_result$L.Ratio[2]` units (p < `r anova_result$"p-value"[2]`), emphasizing its importance as a predictor. Specifically, compared to the Engineering and Technology field, being in the Medical and Health Sciences field is associated with a decrease in RII by `r  coef(model_summary)["fieldMedical and Health Sciences", 1]` units, on average (p = `r coef(model_summary)["fieldMedical and Health Sciences", 4]`), while being in the Social Sciences field is associated with an increase in RII by `r coef(model_summary)["fieldSocial Sciences", 1]` units (p = `r coef(model_summary)["fieldSocial Sciences", 4]`). There are no statistically significant differences between Natural Sciences and Engineering and Technology. Humanities and Agricultural sciences categories were excluded because they had too few observations.
The interaction between treatment and field was not statistically significant, suggesting that the effect of treatment on independence was consistent across disciplines. The three-way interaction among treatment, independence_timing, and field was not statistically significant. This implies that the interaction effect of treatment and independence_timing on independence remained consistent across different disciplines.

<!-- POZN ON MODEL CHOICE: -->

<!-- In Conclusion: -->

<!-- Both the mixed-effects model and the RM-ANOVA suggest a significant interaction between the treatment and the timing of the intervention. This means that the effect of the intervention differed between the treatment and control groups. -->

<!-- Both methods also indicate that the timing of the intervention (i.e., whether it's before or after) has a significant effect on the outcome RII. -->

<!-- The models are in agreement with their primary findings, which is a good sign. It suggests that the treatment has a distinct effect on RII after the intervention, different from that of the control group. -->

<!-- However, it's worth noting that while the methods reach similar conclusions, they approach the problem from different angles. The mixed-effects model incorporates the correlation within each subclass through its random effect, while the RM-ANOVA primarily focuses on the repeated measures nature of the data. Both models provide valuable insights, but the mixed-effects model might be a more robust choice for this particular dataset, given its explicit modeling of the subclass structure. -->

<!-- My take: -->

<!-- The advantage of ANOVa is that it provides information about the "average" effect of independence timing across the dataset, whereas LMM does not do that (but it could be calculated extra, by default it only shows change in time in the control group, not the treatment group, which makes the interpretation harder). On the other hand  -->
<!-- None of these models work well with the disciplines by default - I could make some extra edits to make ANOVA work, but I'm unsure what the interpretation would be for LMM -->

```{r echo=FALSE}

### visualisations

means <- aggregate(RII ~ treatment + independence_timing, data = filtered_by_subclass_1, FUN = mean)

 # ggplot(data = filtered_by_subclass_1, aes(x = treatment, y = RII, fill = independence_timing)) +
 #         geom_violin(scale = "width") +
 #         scale_fill_discrete(name = "Independence Timing", labels = c("Before intervention", "After intervention")) +
 #         xlab("Treatment Group") +
 #         ylab("RII Score") +
 #         ggtitle("Violin Plot of RII Scores by Treatment and Time") +
 #         theme_bw()

 # Calculate means and standard errors for plotting
plot_data <- filtered_by_subclass_1 %>%
  group_by(treatment, independence_timing) %>%
  summarize(
    mean_RII = mean(RII),
    se_RII = sd(RII) / sqrt(n())
  )

# # Plot
# p <- ggplot(plot_data, aes(x = independence_timing, y = mean_RII, group = treatment, color = factor(treatment))) +
#   geom_line(aes(linetype = factor(treatment))) + 
#   geom_point(size = 3) + 
#   geom_errorbar(aes(ymin = mean_RII - se_RII, ymax = mean_RII + se_RII), width = 0.2) +
#   labs(
#     title = "Effect of Intervention on Researcher Independence Index (RII)",
#     x = "Independence Timing",
#     y = "Mean RII",
#     color = "Treatment Group",
#     linetype = "Treatment Group"
#   ) +
#   scale_color_manual(values = c("blue", "red"), labels = c("Control", "Treatment")) +
#   scale_linetype_manual(values = c("dashed", "solid")) +
#   theme_minimal()

dodge_width <- 0.5

combined_plot <- ggplot() +

  geom_violin(data = filtered_by_subclass_1, 
              aes(x = treatment, y = RII, fill = independence_timing), 
              scale = "width", 
              position = position_dodge(dodge_width)) +

  geom_point(data = plot_data, 
             aes(x = treatment, y = mean_RII, color = independence_timing), 
             position = position_dodge(dodge_width), 
             size = 3) +

  geom_errorbar(data = plot_data, 
                aes(x = treatment, ymin = mean_RII - se_RII, ymax = mean_RII + se_RII, color = independence_timing), 
                position = position_dodge(dodge_width), 
                width = 0.2) +

  scale_fill_discrete(name = "Independence Timing", labels = c("Before intervention", "After intervention")) +
  scale_color_manual(values = c("black", "red"), name = "Independence Timing", labels = c("Before intervention", "After intervention")) +

  labs(title = "Violin Plot of RII Scores by Treatment and Time with Mean and CI Overlayed",
       x = "Treatment Group",
       y = "RII Score") +
  theme_bw()

print(combined_plot)

```

```{r echo=FALSE}

# means_disc <- aggregate(RII ~ field, data = filtered_by_subclass_1, FUN = mean)
#      
# num_disc <- filtered_by_subclass_1 %>%
#   group_by(field) %>%
#   summarise(count = n(), .groups = "drop")
# 
# means_disc <- left_join(means_disc, num_disc, by = "field") 
# 
# # Step 1: Reorder the levels of 'field'
# means_disc$field <- with(means_disc, factor(field, levels=unique(field[order(RII)])))
# 
# # Step 2: Optionally, sort the dataframe based on the reordered levels
# means_disc <- means_disc %>% arrange(field)
# 
#     
# se_ci_data <- filtered_by_subclass_1 %>%
#   group_by(field) %>%
#   summarise(mean_RII = mean(RII),
#             se_RII = sd(RII) / sqrt(n()),
#             lower_CI = mean_RII - (1.96 * se_RII),
#             upper_CI = mean_RII + (1.96 * se_RII), 
#             .groups = "drop")
# 
# means_disc <- merge(means_disc, se_ci_data, by = "field")
# 
# 
# ggplot(data = means_disc, aes(x = field, y = mean_RII, fill = field)) +
#   geom_bar(stat = "identity", position = "dodge") +
#   
#   # Add error bars for confidence intervals
#   geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), position = position_dodge(width=0.9), width=0.25) +
#   
#   # Add the number of observations in the middle of the bars
#   geom_text(aes(label=count, y = mean_RII/2), position=position_dodge(width=0.9), vjust=0.5) +
#   
#   # Change x-axis labels and add a title
#   scale_x_discrete(labels = seq_along(unique(means_disc$field))) + # replace with numbers
#   labs(
#     x = "Field", 
#     y = "Mean RII",
#     title = "Differences in RII across disciplines" # Replace with your desired title
#   ) +
#   scale_fill_discrete(name = "Field") +
#   theme_bw()

```

```{r echo=FALSE}

filtered_by_subclass_2 <- pp %>% 
    group_by(subclass) %>% 
    filter(!any(treatment == 0)) %>% 
    ungroup() %>%
    mutate(treatment = as.numeric(as.character(treatment))) %>%
    mutate(treatment = if_else(treatment == 2, 0, treatment)) %>%
    mutate(treatment = factor(treatment, levels = c(0, 1)))

filtered_by_subclass_2$subclass <- as.factor(filtered_by_subclass_2$subclass)
filtered_by_subclass_2$treatment <- as.factor(filtered_by_subclass_2$treatment)
filtered_by_subclass_2$independence_timing <- as.factor(filtered_by_subclass_2$independence_timing)
filtered_by_subclass_2$field <- as.factor(filtered_by_subclass_2$field)

# this DiD model unfortunately doesnt work for some reason that I cannot resolve, but it is possible to calculate the same using repeated measures anova or mixes effects model, so it should be fine
#---> maybe if I edit the dataset so that there is only one row for each observation and it has columns RII_before and RII_after? ------or alternatively I can also just calculate the difference directly RII_diff = after-before and then test only this difference? Not sure how much of a difference this would make -> I should perhaps ask chatGPT

# # Run regression model with interaction term
# did_model <- plm::plm(RII ~ independence_timing * treatment + factor(subclass),
#                  data = filtered_by_subclass_1,
#                  index = c("subclass"),
#                  model = "within")
# 
# # Calculate clustered standard errors
# clustered_se <- coeftest(did_model, vcov = function(x) vcovHC(x, type = "HC1", cluster = "subclass"))

# model2 <- lmer(RII ~ treatment * independence_timing + (1 | subclass), data = filtered_by_subclass_2)
# summary(model2)
# r.squaredGLMM(model2)

result2 <- aov(RII ~ treatment * independence_timing + Error(subclass/independence_timing), data=filtered_by_subclass_2)
# summary(result)
eta_sq_result2 <- eta_squared(result2)

# result2 <- ezANOVA(data = filtered_by_subclass_2,
#                  dv = RII,
#                  wid = subclass,
#                  within = .(independence_timing),
#                  between = treatment)
# print(result)

## adding disciplines:

# result_disc2 <- ezANOVA(data = filtered_by_subclass_2,
#                  dv = RII,
#                  wid = subclass,
#                  within = .(independence_timing),
#                  between = .(treatment, field))
# print(result_disc2)

filtered_data_disc_2 <- filtered_by_subclass_2 %>%
  group_by(field) %>%
  filter(n() >= 10)

filtered_data_disc_2$field <- droplevels(filtered_data_disc_2$field)

# Calculate weights based on group size (you can adjust this based on other criteria if needed)
group_sizes2 <- table(filtered_data_disc_2$field)
max_size2 <- max(group_sizes2)
filtered_data_disc_2$weights <- max_size2 / group_sizes2[filtered_data_disc_2$field]

#this model addresses both unequal group sizes (via weights) and heteroscedasticity (via varIdent). Additionally, I've added a first-order autoregressive structure (corAR1) assuming you have repeated measures on the same subjects and to account for their paired nature (indicated by subclass/independence_timing)
#this model directly tests whether there is a difference in disciplines - anyway it tests each discipline agains each other, so it doesnt produce the neat result of "yes, there are some significant differences between disciplines", but more nuanced version of "there is specifically difference between this and this discipline"

model_gls_weighted2 <- gls(RII ~ field, data = filtered_data_disc_2, 
                          weights = varIdent(form = ~ 1 | field),
                          method = "ML",
                          correlation = corAR1(form = ~ 1 | subject_id))

# summary(model_gls_weighted2)

#this model below is the same as the above but with added predictors

model_gls_interactions_subclass2 <- gls(RII ~ field * treatment * independence_timing, 
                              data = filtered_data_disc_2, 
                              weights = varIdent(form = ~ 1 | field),
                              method = "ML",
                              correlation = corAR1(form = ~ 1 | subclass/independence_timing))

# aa <- summary(model_gls_interactions_subclass2)

#one potential strategy to get the neat result of "yes, there are some significant differences between disciplines" is following: create two similar models

model_without_field2 <- gls(RII ~ treatment + independence_timing + treatment:independence_timing, 
                           data = filtered_data_disc_2, 
                           correlation = corCompSymm(form = ~ 1 | subclass/independence_timing), 
                           method = "ML")

model_with_field_main_effect2 <- gls(RII ~ treatment + field + independence_timing + treatment:independence_timing, 
                                    data = filtered_data_disc_2, 
                                    correlation = corCompSymm(form = ~ 1 | subclass/independence_timing), 
                                    method = "ML")

anova_result2 <- anova(model_without_field2, model_with_field_main_effect2)

model_summary2 <- summary(model_with_field_main_effect2)

```

In the second analysis, we examined the effects of treatment, independence_timing, and their interaction on the researcher independence (RII) for the treatment group and funded control group. Again, there was a significant main effect of time on independence, F(`r format(summary(result2)$"Error: subclass:independence_timing"[[1]]$"Df"[1])`, `r format(summary(result2)$"Error: subclass:independence_timing"[[1]]$"Df"[2])`) = `r format(summary(result2)$"Error: subclass:independence_timing"[[1]]$"F value"[1], digits=2)`, p = `r format(summary(result2)$"Error: subclass:independence_timing"[[1]]$"Pr(>F)"[1], digits=4)`). The effect accounted for approximately `r format(eta_sq_result2[eta_sq_result2$Parameter == "independence_timing", "Eta2_partial"] * 100, digits=2)`% of the observed variance (partial \(\eta^2 = `r eta_sq_result2[eta_sq_result2$Parameter == "independence_timing", "Eta2_partial"]`\), 95% CI [`r eta_sq_result2[eta_sq_result2$Parameter == "independence_timing", "CI_low"]`, `r eta_sq_result2[eta_sq_result2$Parameter == "independence_timing", "CI_high"]`]).

Importantly, this time the interaction effect between treatment and time was not significant, F(`r format(summary(result2)$"Error: Within"[[1]]$"Df"[2])`, `r format(summary(result2)$"Error: Within"[[1]]$"Df"[3])`) = `r format(summary(result2)$"Error: Within"[[1]]$"F value"[2], digits=2)`, p = `r format(summary(result2)$"Error: Within"[[1]]$"Pr(>F)"[2], digits=2)`. This interaction effect explained about `r format(eta_sq_result2[eta_sq_result2$Parameter == "treatment:independence_timing", "Eta2_partial"] * 100, digits=2)`% of the variability (partial \(\eta^2 = `r eta_sq_result2[eta_sq_result2$Parameter == "treatment:independence_timing", "Eta2_partial"]`\), 95% CI [`r eta_sq_result2[eta_sq_result2$Parameter == "treatment:independence_timing", "CI_low"]`, `r eta_sq_result2[eta_sq_result2$Parameter == "treatment:independence_timing", "CI_high"]`]), suggesting that funding has the same effects, regardless of its type or source.

Same as in previous chapter, the comparison of models suggested that there are statistically significant differences in the RII based on the field of study. Specifically, adding the field of study to the model resulted in a significant increase in the log-likelihood by `r anova_result2$L.Ratio[2]` units (p < `r anova_result2$"p-value"[2]`), emphasizing its importance as a predictor. Specifically, compared to the Engineering and Technology field, being in the Medical and Health Sciences field is associated with a decrease in RII by `r  coef(model_summary2)["fieldMedical and Health Sciences", 1]` units, on average (p = `r coef(model_summary2)["fieldMedical and Health Sciences", 4]`), while being in the Social Sciences field is associated with an increase in RII by `r coef(model_summary2)["fieldSocial Sciences", 1]` units (p = `r coef(model_summary2)["fieldSocial Sciences", 4]`). There are no statistically significant differences between Natural Sciences and Engineering and Technology. Humanities and Agricultural sciences categories were excluded because they had too few observations.
The interaction between treatment and field was not statistically significant, suggesting that the effect of treatment on independence was consistent across disciplines. The three-way interaction among treatment, independence_timing, and field was not statistically significant. This implies that the interaction effect of treatment and independence_timing on independence remained consistent across different disciplines.

```{r echo=FALSE}

### visualisations

plot_data2 <- filtered_by_subclass_2 %>%
  group_by(treatment, independence_timing) %>%
  summarize(
    mean_RII = mean(RII),
    se_RII = sd(RII) / sqrt(n())
  )

dodge_width <- 0.5

combined_plot2 <- ggplot() +

  geom_violin(data = filtered_by_subclass_2, 
              aes(x = treatment, y = RII, fill = independence_timing), 
              scale = "width", 
              position = position_dodge(dodge_width)) +

  geom_point(data = plot_data2, 
             aes(x = treatment, y = mean_RII, color = independence_timing), 
             position = position_dodge(dodge_width), 
             size = 3) +

  geom_errorbar(data = plot_data2, 
                aes(x = treatment, ymin = mean_RII - se_RII, ymax = mean_RII + se_RII, color = independence_timing), 
                position = position_dodge(dodge_width), 
                width = 0.2) +

  scale_fill_discrete(name = "Independence Timing", labels = c("Before intervention", "After intervention")) +
  scale_color_manual(values = c("black", "red"), name = "Independence Timing", labels = c("Before intervention", "After intervention")) +

  labs(title = "Violin Plot of RII Scores by Treatment and Time with Mean and CI Overlayed",
       x = "Treatment Group",
       y = "RII Score") +
  theme_bw()

print(combined_plot2)

# means2 <- aggregate(RII ~ treatment + independence_timing, data = filtered_by_subclass_2, FUN = mean)
# 
# 
# means_disc2 <- aggregate(RII ~ field, data = filtered_by_subclass_2, FUN = mean)
#      
# num_disc2 <- filtered_by_subclass_2 %>%
#   group_by(field) %>%
#   summarise(count = n(), .groups = "drop")
# 
# means_disc2 <- left_join(means_disc2, num_disc2, by = "field") 
# 
# # Step 1: Reorder the levels of 'field'
# means_disc2$field <- with(means_disc2, factor(field, levels=unique(field[order(RII)])))
# 
# # Step 2: Optionally, sort the dataframe based on the reordered levels
# means_disc2 <- means_disc2 %>% arrange(field)
# 
# 
# 
# se_ci_data2 <- filtered_by_subclass_2 %>%
#   group_by(field) %>%
#   summarise(mean_RII = mean(RII),
#             se_RII = sd(RII) / sqrt(n()),
#             lower_CI = mean_RII - (1.96 * se_RII),
#             upper_CI = mean_RII + (1.96 * se_RII), 
#             .groups = "drop")
# 
# means_disc2 <- merge(means_disc2, se_ci_data2, by = "field")
# 
# 
# ggplot(data = means_disc2, aes(x = field, y = mean_RII, fill = field)) +
#   geom_bar(stat = "identity", position = "dodge") +
#   
#   # Add error bars for confidence intervals
#   geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), position = position_dodge(width=0.9), width=0.25) +
#   
#   # Add the number of observations in the middle of the bars
#   geom_text(aes(label=count, y = mean_RII/2), position=position_dodge(width=0.9), vjust=0.5) +
#   
#   # Change x-axis labels and add a title
#   scale_x_discrete(labels = seq_along(unique(means_disc2$field))) + # replace with numbers
#   labs(
#     x = "Field", 
#     y = "Mean RII",
#     title = "Differences in RII across disciplines" # Replace with your desired title
#   ) +
#   scale_fill_discrete(name = "Field") +
#   theme_bw()

# disc_data <- pp %>% 
#     filter(independence_timing != "after_intervention")

disc_data <- d %>%
     filter(!is.na(RII)) %>%
     group_by(vedidk, field, RII) %>%
     sample_n(1) %>%
     ungroup()

means_disc_complete <- aggregate(RII ~ field, data = disc_data, FUN = mean)

num_disc_complete <- disc_data %>%
  group_by(field) %>%
  summarise(count = n(), .groups = "drop")

means_disc_complete <- left_join(means_disc_complete, num_disc_complete, by = "field")

# Step 1: Reorder the levels of 'field'
means_disc_complete$field <- with(means_disc_complete, factor(field, levels=unique(field[order(RII)])))

# Step 2: Optionally, sort the dataframe based on the reordered levels
means_disc_complete <- means_disc_complete %>% arrange(field)

se_ci_data_complete <- disc_data %>%
  group_by(field) %>%
  summarise(mean_RII = mean(RII),
            se_RII = sd(RII) / sqrt(n()),
            lower_CI = mean_RII - (1.96 * se_RII),
            upper_CI = mean_RII + (1.96 * se_RII), 
            .groups = "drop")

means_disc_complete <- merge(means_disc_complete, se_ci_data_complete, by = "field")

ggplot(data = means_disc_complete, aes(x = field, y = mean_RII, fill = field)) +
  geom_bar(stat = "identity", position = "dodge") +

  # Add error bars for confidence intervals
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), position = position_dodge(width=0.9), width=0.25) +

  # Add the number of observations in the middle of the bars
  geom_text(aes(label=count, y = mean_RII/2), position=position_dodge(width=0.9), vjust=0.5) +

  # Change x-axis labels and add a title
  scale_x_discrete(labels = seq_along(unique(means_disc_complete$field))) + # replace with numbers
  labs(
    x = "Field", 
    y = "Mean RII",
    title = "Differences in average RII across disciplines"
  ) +
  scale_fill_discrete(name = "Field") +
  theme_bw()

```

# Discussion

The results of this study provide important insights into the effects of funding interventions on researcher independence. The strongest effect we found across all control group comparison was the effect of time - over time, researchers become more independent, regardless of their access to funding or their study discipline. This effect explains around 60% of the variance in independence.

Additionally, the observed variation in independence across different disciplines underscores the importance of considering disciplinary differences, suggested by a moderate effect size. A moderate effect size for the field effect emphasizes that certain disciplines might inherently have variations in RII, perhaps due to differences in practices, methodologies, or norms specific to each discipline. Anyway, these differences remain constant across time and funding interventions, suggesting that funding has a uniform effect on the independence across disciplines.

Finally, the significant interaction between treatment and time when comparing treatment vs unfunded group suggests the funding has positive effect on stimulating researchers' independence. However, the effect size is only modest, explaining about 2 % of variance in independence, indicating relatively small practical relevance. Further, the insignificant interaction when comparing the treatment vs funded control group suggests that funding in this career stage has the same effect, regardless of its type or source. Interestingly, in our sample both funded groups (treatment as well as control) started with lower independence compared to unfunded control group. This might suggest that in the early career stages, in order to secure funding, it is needed to be more dependent on one's PhD advisor.

<!-- Anyway, rather than granting funded researchers greater independence compared to the unfunded control group, it seems that the funding programme we examined sampled less independent researchers and then allowed them to catch up with the unfunded control group via providing them funding. -->

## Limitations

--> only talks about the Czech environment and considers researchers growing up and most of their careers in the Czech republic

# Attachments

## Identification of supervisors

Data about PhD student-PhD advisor connections are usually not available. One must search for these connections manually on the internet, which is time consuming and not fully reliable (we were only able to find about 69 % of these connections in our treatment group). Therefore, we tried to come up with an alternative way of identifying supervisors from the available data: we assumed that since many researchers only start their publication career during their PhD studies, their PhD supervisors might often coauthor their first publications. That led us to creating an algorithm to identify the supervisor solely from the student's publication history as "the most frequent co-author of the first 5 publications".

To check whether we can use "the most frequent co-author of the first 5 publications" as a proxy to identify supervisors, we have compared the supervisors identified via this method with the manually found supervisors for each researcher in our sample.
In the treatment group of `r nrow(tar_read(ids_complete))`researchers,  118 (49 %) had their most frequent co-author same as their supervisor and 86 (36 %) had someone other than their supervisor as their most frequent co-author in their first 5 publications, and 37 (15%) had no co-author in their first 5 publications.  
In the control groups, we have manually found supervisors for `r nrow(tar_read(sup_complete) %>% filter(!is.na(sup_vedidk)) %>% distinct(name_first, name_last))` researchers, and when we compared them to the supervisors suggested by the most frequent coauthor algorithm, we found that 41 (29 %) had their most frequent co-author the same as their supervisor, 80 (57 %) had someone other than their supervisor as their most frequent co-author in their first 5 publications, and 20 (14 %) had no co-author in their first 5 publications.

Further, we have conducted analysis on an independent sample of researchers from Slovenia (citace Luzar??), and found that  when we filter out all researchers who have not had any publication 4 years prior to their PhD defense (1135, approx 9,5% of the sample), we see that from the remaining researchers, about 50,3 % had their supervisor as the most frequent coauthor in the 4 years prior to their PhD defense.

# References